Binary companions of red giant stars in \apogee\ \DR

### 2018-03-04 paper/1-catalog/twoface-catalog.tex
MAP not median

% define macros for text
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\sdssiv}{\project{\acronym{SDSS-IV}}}
\newcommand{\thejoker}{\project{The~Joker}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\DR}{\acronym{DR14}}
\newcommand{\RC}{\acronym{RC}}
\newcommand{\RGB}{\acronym{RGB}}

\newcommand{\nprior}{536,870,912}
\newcommand{\nposterior}{256}
\newcommand{\nstars}{96,231}
\newcommand{\nvisits}{397,559}
\newcommand{\ncontrol}{16,384}
\newcommand{\nhighK}{4,510}
\newcommand{\nunimodal}{XX}
\newcommand{\nmultimodal}{XX}

@@

We also run on the null control sample (see \sectionname~\ref{sec:control}) with
the same parameters.
From the control sample run, 13 ``stars'' are flagged as \emph{needs MCMC}, and
1023 are flagged as ``needs more prior samples'';
For these stars, we don't continue sampling with \thejoker\ or MCMC and only use
the posterior samples returned from the \thejoker.
% See: Sample K cuts.ipynb

@@

We don't expect there to be a sharp transition in orbital parameters between
stars with companions and stars without companions: There is a continuum of
companion properties.
For example, the companion masses can be low mass, or at long periods, or at
high inclination, all of which will make the velocity semi-amplitude, $K$, small
for a given system.
Therefore, from the radial velocity data alone, there is no way to define a
simple cut to select a complete sample of stars with companions.
It is nevertheless possible to define a cut that selects stars with
high-confidence companions.

We select a sample of stars that confidently have companions using percentiles
computed from the posterior samples in the velocity semi-amplitude parameter,
$K$.
\figurename~\ref{fig:lnK-percentiles}, shows the distribution of 1st percentiles
of $\ln K$ computed for all stars with posterior samplings from \thejoker\
(filled, blue histogram), along with the same for posterior samplings for the
null control sample (solid, dark line).
To select stars with companions, we use a cut in the 1st percentile of the $\ln
K$ samples such that $<1\%$ of the control sample is selected, i.e. our
estimated false positive rate is $<1\%$.
We use a threshold of $\ln K = 0$ to meet this constraint (vertical, dashed line
in \figurename~\ref{fig:lnK-percentiles}); \nhighK\ stars pass this cut.
For brevity below, we refer to this sample as the ``high-$K$'' stars, and the
complimentary sample as the ``low-$K$'' stars.

@@

\begin{table}
    \centering
    \begin{tabular}{c | c}
    \texttt{APOGEE\_ID} & \texttt{lnK\_per\_1} \\
    \hline
    2M00000002+7417074 & -2.028 \\
    2M00000068+5710233 & -2.600 \\
    2M00000222+5625359 &  2.134 \\
    2M00000446+5854329 & -7.199 \\
    2M00000535+1504343 & -4.457 \\
    2M00000546+6152107 & -4.388 \\
    2M00000797+6436119 & -0.774 \\
    2M00000866+7122144 & -3.313 \\
    ... & ...
    \end{tabular}
    \caption{This table contains the \texttt{APOGEE\_ID} for all \nstars\ stars
    in the parent sample used in this work, for which we have posterior
    samplings in orbital parameters.
    The other column contains the 1st percentile computed over the $\ln K$
    samples for each source, \texttt{lnK\_per\_1}:
    The ``high-$K$'' sample (see \sectionname~\ref{sec:catalog-confident}) is
    defined using this column, by selecting \texttt{lnK\_per\_1 > 0}.
    }
    \label{tbl:lnK-per}
\end{table}

\tablename~\ref{tbl:lnK-per} contains a list of all stars in the parent sample
and the value of the 1st percentile over the $\ln K$ samples for each star.
\figurename s~\ref{fig:highK-0} and \ref{fig:highK-1} show eight examples of
high-$K$ stars, i.e. stars that likely have companions, with different numbers
of visits, $N$, indicated on the panels.
\figurename s~\ref{fig:lowK-0} and \ref{fig:lowK-1} is the same for eight
low-$K$ stars, i.e. stars that either have no companions, low-mass companions,
or companions at long periods or high inclination.

The majority of the stars in the high-$K$ sample have poorly constrained orbital
parameters because the posterior pdfs are very multimodal.
The high-$K$ sample includes all stars in the catalogs described in the next two
subsections, however these two catalogs are mutually exclusive with one another.

@@

A subset of the high-$K$ sample (\sectionname~\ref{sec:catalog-confident}) have
multimodal posterior distributions over orbital parameters that are limited to
just a few qualitatively different solutions.
For these stars, one or a few more well-timed RV measurements would likely lead
to uniquely-determined orbital parameters.
\todo{How exactly do we select these stars?}

\figurename~\ref{fig:todo} shows a few examples of stars that meet these
criteria.
Note that in these cases, there are certain times at which a future observation
would be far more informative: For example, \todo{top panel, time X vs. time Y}.

@@

\todo{Describe catalog contents: Also include M2min and primary mass when
possible?}

@@

\todo{Also include M2min and primary mass when possible?}

@@

>>>>

% define macros for text
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\sdssiv}{\project{\acronym{SDSS-IV}}}
\newcommand{\thejoker}{\project{The~Joker}}
\newcommand{\thecannon}{\project{The~Cannon}}
\newcommand{\DR}{\acronym{DR14}}
\newcommand{\RC}{\acronym{RC}}
\newcommand{\RGB}{\acronym{RGB}}
\newcommand{\quantile}[2]{\ensuremath{\textrm{Q}_{#1}}\left(#2\right)}

\newcommand{\nprior}{536,870,912}
\newcommand{\nposterior}{256}
\newcommand{\nstars}{96,231}
\newcommand{\nvisits}{397,559}
\newcommand{\ncontrol}{16,384}
\newcommand{\nhighK}{4,510}
\newcommand{\nmultimodal}{XX}
\newcommand{\nunimodal}{319}

@@

We also run on the null control sample (see
\sectionname~\ref{sec:control-sample}) with the same parameters.
From the control sample run, 13 ``stars'' are flagged as \emph{needs MCMC}, and
1023 are flagged as ``needs more prior samples'';
For these control sample stars, we don't continue sampling with \thejoker\ or
MCMC and only use the ($<$\nposterior) posterior samples returned from the
\thejoker.
% See: Sample K cuts.ipynb

@@

We don't expect there to be a sharp transition in orbital parameters between
stars with companions and stars without companions: There is a continuum of
companion properties.
For example, the companions can be low mass, or at long periods, or at high
inclination, all of which will make the velocity semi-amplitude, $K$, small for
a given system.
Therefore, from the radial velocity data alone, there is no way to define a
simple cut to select a complete sample of stars with companions.
It is nevertheless possible to define a cut that selects stars with
high-confidence companions.

We select a sample of stars that confidently have companions using percentiles
computed from the log of the posterior samples in the velocity semi-amplitude
parameter, $\ln K$.
\figurename~\ref{fig:lnK-percentiles}, shows the distribution of 1st percentiles
of $\ln K$ computed for all stars with posterior samplings from \thejoker\
(filled, blue histogram), along with the same for posterior samplings for the
null control sample (solid, dark line).
To select stars with companions, we use a cut in the 1st percentile of the $\ln
K$ samples such that $<1\%$ of the control sample is selected, i.e. our
estimated false positive rate is $<1\%$.
We use a threshold of $\ln K = 0$ to meet this constraint (vertical, dashed line
in \figurename~\ref{fig:lnK-percentiles}); \nhighK\ stars pass this cut.
For brevity below, we refer to this sample as the ``high-$K$'' stars, and the
complimentary sample as the ``low-$K$'' stars.

@@

\begin{table}
    \centering
    \begin{tabular}{c | c}
    \texttt{APOGEE\_ID} & \texttt{lnK\_per\_1} \\
    \hline
    2M00000002+7417074 & -2.028 \\
    2M00000068+5710233 & -2.600 \\
    2M00000222+5625359 &  2.134 \\
    2M00000446+5854329 & -7.199 \\
    2M00000535+1504343 & -4.457 \\
    2M00000546+6152107 & -4.388 \\
    2M00000797+6436119 & -0.774 \\
    2M00000866+7122144 & -3.313 \\
    ... & ...
    \end{tabular}
    \caption{This table contains \texttt{APOGEE\_ID}'s for all \nstars\ stars
    in the parent sample used in this work, for which we have posterior
    samplings in orbital parameters.
    The other column contains the 1st percentile computed over the $\ln K$
    samples for each source, \texttt{lnK\_per\_1}:
    The ``high-$K$'' sample (see \sectionname~\ref{sec:catalog-confident}) is
    defined using this column, by selecting \texttt{lnK\_per\_1 > 0}.
    }
    \label{tbl:lnK-per}
\end{table}

\tablename~\ref{tbl:lnK-per} contains a list of all stars in the parent sample
and the value of the 1st percentile over the $\ln K$ samples for each star.
\figurename s~\ref{fig:highK-0} and \ref{fig:highK-1} show eight examples of
high-$K$ stars, i.e. stars that likely have companions, with different numbers
of visits, $N$, indicated on the panels.
\figurename s~\ref{fig:lowK-0} and \ref{fig:lowK-1} shows the same for eight
low-$K$ stars, i.e. stars that either have no companions, low-mass companions,
or companions at long periods or high inclination.
The majority of the stars in the high-$K$ sample have poorly constrained orbital
parameters because the posterior pdfs are very multimodal.
The high-$K$ sample selected from \tablename~\ref{tbl:lnK-per} includes all
stars in the catalogs described in the next two subsections, however the
following two catalogs are mutually exclusive.

@@

A subset of the high-$K$ sample (\sectionname~\ref{sec:catalog-confident}) have
multimodal posterior distributions over orbital parameters that are limited to
a few qualitatively different solutions.
For these stars, one or a few more well-timed RV measurements would likely lead
to uniquely-determined orbital parameters.
\figurename~\ref{fig:todo} shows a few examples of stars that meet these
criteria.
These stars are prime candidates for follow-up efforts.
In many of these cases, there are certain times at which a future observation
would be far more informative.
Visually, these times correspond to regions of time when the predicted RVs have
large variance as computed for orbits computed from the posterior samplings.
For example, \todo{top panel, time X vs. time Y, node vs. large variance}.

\tablename~\ref{tbl:todo} contains ...
\todo{Describe catalog contents: Also include M2min and primary mass when
possible?}

@@

@@

Starting from the high-$K$ sample, we select stars that have posterior period
samples that fall within a single mode.
We define a period resolution for each star $\Delta =
\left[4\,P_{\textrm{min}}^2 / (2\pi \, T) \right]$, where $P_{\textrm{min}}$ is
the minimum period sample value, and $T$ is the epoch span of the data for that
star.
We consider a sampling to be unimodal when $P_{\textrm{max}} - P_{\textrm{min}}
< \Delta$ (see \sectionname~2 of \citealt{Price-Whelan:2017}).
Unlike the multi-modal stars, the posterior pdfs for these stars can be
approximated using point-estimates and standard deviations of their respective
samplings.

\tablename~\ref{tbl:highK-unimodal} contains ...

\input{tables/unimodal-schema}

@@

\section{Using data products...}
\label{sec:code-demo}

\todo{Brief example of how to use twobody and thejoker to visualize orbits, with
link to a tutorial notebook on github}
